The cost of undisturbed landscapes

### 2020-04-21 projects/asparagus/doc/slides.tex

\documentclass[aspectratio=169, xcolor=dvipsnames]{beamer}
\usetheme{Pittsburgh}
\usepackage{tikz}
\usepackage{booktabs}
\usepackage[version=4]{mhchem}
\usepackage{animate}
\usepackage{biblatex}
\usepackage[absolute,overlay]{textpos}
\usepackage{fancyvrb}
\usepackage{xcolor}
\usepackage{multicol}
\usepackage{setspace}
\usepackage{siunitx}
\DeclareSIUnit\year{a}
\usepackage{qrcode}
\usepackage{eurosym}
\usepackage{xfrac}

@@

\setbeamertemplate{section page}{
	\begin{textblock*}{4cm}(11cm, 4cm) % {block width} (coords)
	{\LARGE \insertsection}
	\end{textblock*}
}

\begin{document}
    %%%%Title Page
    {
        \usebackgroundtemplate{
	        \begin{picture}(320,300)
    	        \hspace{9.0cm}
	              \includegraphics[width=0.7\linewidth]{./logos/refuel_logo_with_text}
            \end{picture}
        }

@@

\section{Introduction}
\begin{frame}
\frametitle{Austrian energy policy objectives}
\framesubtitle{according to government programme 2020-2024}
\begin{itemize}
\item 100\%\footnote[frame]{excluding system services and industry own consumption. At current levels this equals 10\% of consumption, i.e. actual target is around 90\%.} of electricity demand from domestic renewable sources on annual balance by 2030 \pause
\item additional annual electricity generation of \SI{27}{\tera\watt\hour} is expected to suffice \pause
\item technology-specific additions:
\end{itemize}

\begin{table}
\centering
\begin{tabular}{| l | r | r | r | r | r |}
\hline
 & 2018 & 2018\footnote[frame]{meteorological conditions as in 2016} & Policy & 2030 & 2030 \\ 
 & [GW] & [TWh] & [TWh] & [TWh] & [GW] \\ \hline \hline
Solar PV & 1.44 & 1.23 & +11 & 12.23 &  14.27 \\ \hline
Wind (onshore) & 3.05 & 6.14 & +10 & 16.14 & 8 \\ \hline
Hydro (run-of-river) & 5.72 & 28.34 & +5 & 33.34 & 6.73 \\ \hline
%Biomass & 1 & 5 & +1 & 6 & 2 \\ \hline
\end{tabular}
\end{table}
\end{frame}

\begin{frame}
\frametitle{Stylized Facts}
\begin{itemize}
\item Apart from wind and solar, potentials for renewable electricity generation in Austria largely exhausted
\item Under announced policies, solar PV is only large-scale substitute to wind power
\item Increasing social conflict around the large-scale expansion of onshore wind power
\item Traditional power system models do not account for local negative externalities of renewable energy generators, such as:
	\begin{itemize}
	\item visual impact on landscape
	\item harm to wildlife
	\item noise, flickering, glaring
	\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Why wind is not solar}
\framesubtitle{A side note on imperfect substitutes}
\begin{textblock*}{10cm}(0.75cm, 2.0cm) % {block width} (coords)
\centering
Aggregate generation profile in Austria
\includegraphics[width=\textwidth, keepaspectratio]{./figures/wind_not_pv}
\end{textblock*}
\begin{textblock*}{4cm}(11cm, 2.25cm)
%Aggregate profile for Austria
\bigskip
\begin{itemize}
\item secure system operation requires $S=D$ at any point in time
\item electricity not easily storable
\item grid operators >>transform<< electricity feed-in to end-use electricity
\end{itemize}
\end{textblock*}
\end{frame}

\begin{frame}
\frametitle{The problem from a social planner's perspective}
How to account for local negative externalities in renewables expansion planning?

@@

\section{Research agenda}
\begin{frame}
\frametitle{The cost of undisturbed landscapes}
\framesubtitle{Approach}
\begin{itemize}
\item Resemble Austrian power system in 2030
\item include most important electricity trading partner Germany
\item set policy target of meeting $90\%$ of demand in 2030 from domestic renewable sources
\item incorporate announced electricity system targets for Germany in 2030
	\begin{itemize}
	\item nuclear phase-out
	\item partial coal exit
	\item expansion of renewable capacities in line with EEG 2017
	\end{itemize}
\item simulate prospective electricity system with \emph{medea}
\end{itemize}
\end{frame}

@@

\begin{column}{0.325\textwidth}
\textbf{Implementation}
\begin{itemize}
\item linear program
\item python \& GAMS
\end{itemize}
\end{column}
\end{columns}
\end{footnotesize}
\medskip
\begin{center}
\emph{medea} is available on \href{https://github.com/inwe-boku/medea}{github.com/inwe-boku/medea} under an open MIT license
\end{center}
\end{frame}
\note{
medea is a partial-equilibrium model of the coupled electricity and district heating sectors.
Operation of and investment in conventional and renewable power plants, storages and transmission lines are determined simultaneously for one year.
Results include hourly dispatch, investment, inter-zonal trade, electricity and heat prices, profits, and emissions of air pollutants.
The model accounts for various technical characteristics of power systems, such as co-generation of heat and power, balancing requirements, 
}

@@

\section{Discussion}
\begin{frame}
\frametitle{The cost of undisturbed landscapes}
\framesubtitle{Discussion of results}
\begin{itemize}
\item Renewable resource quality held constant as capacity is expanded
\item Sub-national electricity transmission and distribution grids neglected
\item Technical operation of generators not fully represented \\(e.g. no unit commit, simplified balancing)
\item Electricity market-splitting has increased market concentration
\item Announced policy necessarily turns Austria into a net exporter of electricity\\
$\rightarrow$ "loop-flows" potentially avoided \\
$\rightarrow$ artificial transmission restriction between DE and AT could be eliminated
%\item use Austrian landscape for "overinvesting" in wind and exporting to Germany?
\end{itemize}
\end{frame}

@@

    \end{frame}
}   
\appendix
\backupbegin

\begin{frame}
\frametitle{The cost of undisturbed landscapes}
\framesubtitle{Unconstrained deployment of wind and solar power}

\begin{textblock*}{10cm}(0.75cm, 1.8cm) % {block width} (coords)
\begin{table}
\begin{tabular}{l r l}
\emph{Results for Austria} \\ \hline \hline
Added Solar PV & 0.0 & \SI{}{\tera\watt\hour} \\ \hline
Added Wind & 31.0 & \SI{}{\tera\watt\hour} \\ \hline \hline
System Cost\footnote[frame]{excluding cost for domestic transmission and distribution grids} & $3095.2$ & \SI{}{\mega\text{\euro}} \\ \hline
Trade balance & $746.8$ & \SI{}{\mega\text{\euro}} \\ \hline
Cost of air pollution & $184.4$ & \SI{}{\mega\text{\euro}} \\ \hline
\ce{CO2} emissions & $5227.3$ & \SI{}{\kilo\tonne} \\ \hline
Electricity generation & $92.8$ & \SI{}{\tera\watt\hour} \\ \hline
Curtailment & $2.4$ & \SI{}{\tera\watt\hour} \\ \hline
Net exports & $5.9$ & \SI{}{\tera\watt\hour} \\ \hline
Wholesale electricity price & $56.05$ & \SI{}{\text{\euro}\per\mega\watt\hour} \\ \hline
\end{tabular}
%\caption{Results for Austria}
\end{table}
\end{textblock*}

\begin{textblock*}{4cm}(11cm, 2.05cm)
\begin{small}
\begin{table}
\begin{tabular}{p{1.8cm} | p{0.3cm} l}
Capital cost of solar PV & $630$ & \SI{}{\text{\euro}\per\kilo\watt\text{p}} \\
\ce{CO2} price & $90$ & \SI{}{\text{\euro}\per\mega\watt\hour}
\end{tabular}
\end{table}
\end{small}
\end{textblock*}

>>>>

\documentclass[aspectratio=1610, xcolor=dvipsnames,handout]{beamer} %aspectratio=169, 
\usetheme{Pittsburgh}
\usepackage{tikz}
\usepackage{booktabs}
\usepackage[version=4]{mhchem}
\usepackage{animate}
\usepackage{biblatex}
\usepackage[absolute,overlay]{textpos}
\usepackage{fancyvrb}
\usepackage{xcolor}
\usepackage{multicol}
\usepackage{setspace}
\usepackage{siunitx}
\DeclareSIUnit\year{a}
\usepackage{qrcode}
\usepackage{eurosym}
\usepackage{xfrac}
\usepackage[export]{adjustbox}

@@

%\setbeamertemplate{section page}{
%	\begin{center}
%	{\usebeamerfont[fg]{section title} \insertsection}
%	\end{center}
%}
\AtBeginSection[]{
  \begin{frame}
  \vfill
%  \centering
%  \begin{beamercolorbox}[sep=8pt,center,shadow=false,rounded=true]{title}
\tableofcontents[currentsection]
%\begin{textblock*}{10cm}(7cm, 2.0cm) % {block width} (coords)
%\adjincludegraphics[height=6cm,trim={0 0 {.3\width} 0},clip]{./figures/pv}
%\end{textblock*}
 %   \usebeamerfont{title} \usebeamercolor{title} \insertsection\par%
%  \end{beamercolorbox}
  \vfill
  \end{frame}
}

\begin{document}
    %%%%Title Page
    {
        \usebackgroundtemplate{
	        \begin{picture}(320,325)
    	        \hspace{9.0cm}
	              \includegraphics[width=0.7\linewidth]{./logos/refuel_logo_with_text}
            \end{picture}
        }

@@

\section{Background}
\begin{frame}
\frametitle{Austrian energy policy objectives}
\framesubtitle{according to government programme 2020-2024}
\begin{itemize}
\item 100\%\footnote[frame]{excluding system services and industry own consumption. At current levels this equals 10\% of consumption, i.e. actual target is around 90\%.} of electricity demand from domestic renewable sources on annual balance by 2030 \pause
\begin{itemize}
\item[$\rightarrow$] necessarily turns Austria into a net exporter of electricity \pause
\end{itemize}
\item additional annual electricity generation of \SI{27}{\tera\watt\hour} is expected to suffice \pause
\item technology-specific additions:
\end{itemize}

\begin{table}
\centering
\begin{tabular}{| l | r | r | r | r | r |}
\hline
 & 2018 & 2018\footnote[frame]{meteorological conditions as in 2016} & Policy & 2030 & 2030 \\ 
 & [GW] & [TWh] & [TWh] & [TWh] & [GW] \\ \hline \hline
Solar PV & 1.44 & 1.23 & +11 & 12.23 &  14.27 \\ \hline
Wind (onshore) & 3.05 & 6.14 & +10 & 16.14 & 8 \\ \hline
Hydro (run-of-river) & 5.72 & 28.34 & +5 & 33.34 & 6.73 \\ \hline
Biomass & $\sim 1$ & 4.78 & +1 & 5.78 & $>2$ \\ \hline
\end{tabular}
\end{table}
\end{frame}

\begin{frame}
\frametitle{Why wind is not solar}
\framesubtitle{A side note on imperfect substitutes}
%\begin{textblock*}{4cm}(0.75cm, 2.75cm)
%Aggregate profile for Austria
\begin{textblock*}{4cm}(0.75cm, 2.75cm)
\begin{itemize}
\item secure system operation requires $S=D$ at any point in time 
\item electricity not easily storable 
\item >>transforming<< electricity feed-in to end-use electricity is costly
\end{itemize}
\end{textblock*}

\only<2>{
\begin{textblock*}{10cm}(5.5cm, 2.25cm) % {block width} (coords)
\centering
Aggregate generation profile in Austria
\includegraphics[width=\textwidth, keepaspectratio]{./figures/wind_not_pv}
\end{textblock*}
}
\end{frame}

\begin{frame}
\frametitle{Stylized Facts}
\begin{itemize}
\item Apart from wind and solar, potentials for renewable electricity generation in Austria largely exhausted
\item Under announced policies, solar PV is only large-scale substitute to wind power \pause
\item Increasing social conflict around the large-scale expansion of onshore wind power
\item Traditional power system models do not account for local negative externalities of renewable energy generators, such as:
	\begin{itemize}
	\item visual impact on landscape
	\item harm to wildlife
	\item noise, flickering, glaring
	\end{itemize}
\end{itemize}
\end{frame}

\section{Motivation \& Method}
\begin{frame}
\frametitle{The problem from a social planner's perspective}
How to account for local negative externalities in renewables expansion planning?

@@

\begin{frame}
\frametitle{The cost of undisturbed landscapes}
\framesubtitle{Approach}
\begin{itemize}
\item Resemble Austrian power system in 2030
\item include most important electricity trading partner Germany \pause
\item set policy target of meeting $90\%$ of demand in 2030 from domestic renewable sources
\item incorporate announced electricity system targets for Germany in 2030
	\begin{itemize}
	\item nuclear phase-out
	\item partial coal exit
	\item expansion of renewable capacities in line with EEG 2017
	\end{itemize} \pause
\item simulate prospective electricity system with \emph{medea}
\end{itemize}
\end{frame}

@@

\begin{column}{0.325\textwidth}
\textbf{Implementation}
\begin{itemize}
\item linear program
\item python \& GAMS
\end{itemize}
\end{column}
\end{columns}
\end{footnotesize}
\bigskip

\medskip
\begin{center}
\emph{medea} is available on \href{https://github.com/inwe-boku/medea}{github.com/inwe-boku/medea} under an open MIT license
\end{center}
\end{frame}
\note{
medea is a partial-equilibrium model of the coupled electricity and district heating sectors.
Operation of and investment in conventional and renewable power plants, storages and transmission lines are determined simultaneously for one year.
Results include hourly dispatch, investment, inter-zonal trade, electricity and heat prices, profits, and emissions of air pollutants.
The model accounts for various technical characteristics of power systems, such as co-generation of heat and power, balancing requirements, 
}

@@

\section{Discussion \& Conclusion}
\begin{frame}
\frametitle{The cost of undisturbed landscapes}
\framesubtitle{Discussion of results}
\begin{itemize}
\item Renewable resource quality held constant as capacity is expanded
\item Sub-national electricity transmission and distribution grids neglected
\item Technical operation of generators not fully represented \\(e.g. no unit commit, simplified balancing)
\item Electricity market-splitting has increased market concentration
\item Announced policy necessarily turns Austria into a net exporter of electricity\\
$\rightarrow$ "loop-flows" potentially avoided \\
$\rightarrow$ artificial transmission restriction between DE and AT could be eliminated
%\item use Austrian landscape for "overinvesting" in wind and exporting to Germany?
\end{itemize}
\end{frame}

@@
